analysis

### 2023-08-03 analysis/paper/data analysis.Rmd
code style

```{r}
library(readr)
# import decor data
decors <- read_csv2(here::here("analysis/data/raw_data/decors_late.csv"))
# import information on sites
sites <- read_csv2(here::here("analysis/data/raw_data/sites_all.csv"))
```

@@

```{r total network sample, fig.cap="Partially overlapping networks of four decorative practices"}
decor_graph %>%
  activate(edges) %>%
  filter(name %in% c("braided_cord_roulette", "braided_strip_roulette", "paint", "straight_cord_mat")) %>%
  ggraph() +
  coloured_edges +
  theme_graph()
```

@@

# calculate modularity of louvain clustering
modularity <- decor_graph %>%
  activate(nodes) %>%
  mutate(modularity = graph_modularity(group = as.factor(louvain))) %>%
  pull(modularity) %>%
  head(1)
```

@@

```{r relative edge weights and centralization by cluster}
# Relative edge weights (how important are the connections within the cluster)
rel_edge_weight <- vector("double", length(3))
for (i in 1:3) {
  rel_edge_weight[[i]] <- decor_graph %>%
    # filter community
    activate(nodes) %>%
    filter(louvain == i) %>%
    # relative edge weight (sum of weights / number of edges)
    activate(edges) %>%
    as.data.frame() %>%
    {
      sum(.$weight) / nrow(.)
    }
}

@@

# put together
tibble(rel_edge_weight, centralization) %>%
  mutate(cluster = 1:3) %>%
  mutate(rel_edge_weight = rel_edge_weight / 100) %>%
  rename("relative edge weight" = rel_edge_weight) %>%
  pivot_longer(!cluster, names_to = "key", values_to = "value") %>%
  # plot
  ggplot(aes(x = cluster, y = value, fill = key)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_grey() +
  theme(axis.text.x = element_text(angle = 90)) +
  theme_publish()
```

@@

  # find the edges that cross between communities
  mutate(crossing = community_crossing(weights = weight)) %>%
  filter(
    crossing == TRUE,
    weight > 200
  ) %>%
  activate(nodes) %>% 
  filter(!node_is_isolated()) %>%

@@

@@

# Simplified network backbone

@@

```{r backbone extraction}
# extract binary backbones at decreasing levels of significance
decor.05 <- extract_backbone(decor.flat, 0.05)
decor.15 <- extract_backbone(decor.flat, 0.15)
decor.20 <- extract_backbone(decor.flat, 0.20)

@@

```{r import early sites}
library(readr)
decorsearly <- read_csv2(here::here("analysis/data/raw_data/decors_early.csv"))

# apply function to transform frequency table into edge list
edgelist <- make_edgelist(decorsearly)

@@

>>>>

```{r}
library(readr)
# import decor data

#decors late is the "cleaned" version without Oudalan and SAO. they are still present in the earlier version decors.csv
decors <- read_csv(here::here("analysis/data/raw_data/decors.csv"))

decors <- decors %>%
  filter(site != "SAO 3" & site != "Oudalan fer recent")  %>%
  mutate_at(vars(-site), as.numeric) %>% 
  mutate(site = case_when(site == "HAMB 4" ~ "Hamb 4",
         site == "KAN 4" ~ "Kaniana 4",
         site == "Akumbu 1" ~ "Akumbu late",
         site == "TMD" ~ "TMD 5",
         TRUE ~ site)) %>% 
  select(-slip)

 decors[is.na(decors)] <- 0

 #write.csv2(decors,here::here("analysis/data/raw_data/decors_late.csv"))

# import information on sites
sites <- read_csv2(here::here("analysis/data/raw_data/sites_all.csv"))
```

@@

```{r total network sample, fig.cap="Partially overlapping networks of four decorative practices"}
decor_graph %>%
  activate(edges) %>%
  filter(name %in% c("braided_cord_roulette", "braided_strip_roulette", "paint", "straight_cord_mat"))%>%
  ggraph() +
  coloured_edges +
  theme_graph()

decor_graph %>%
  activate(edges) %>%
  filter(name %in% c("folded_strip_roulette", "braided_strip_roulette", "paint", "straight_cord_mat", "twisted_cord_roulette", "braided_cord_roulette"))%>%
  filter(weight >= 3) |> 
  activate(nodes) |> 
  filter(!node_is_isolated()) |> 
  ggraph() +
  coloured_edges +
  theme_graph()

decor_graph %>%
  activate(edges) %>%
  filter(name %in% c("folded_strip_roulette","braided_strip_roulette", "paint", "straight_cord_mat", "braided_cord_roulette"))%>%
  filter(weight >= 3) |> 
  activate(nodes) |> 
  filter(!node_is_isolated()) |> 
  ggraph() +
  coloured_edges +
  theme_graph()

decor_graph |> 
    activate(edges) |> 
  # Filters 'edges' based on 4 specific decorations
  filter(name %in% c('folded_strip_roulette', 'braided_strip_roulette', 'paint', 'straight_cord_mat')) |> 
#  filter(weight > 1) |> 
  #plot the graph
    ggraph(layout = 'stress') + 
    coloured_edges + 
    theme_graph() 
```

@@

# calculate modularity of louvain clustering
modularity <- decor_graph %>%
  activate(nodes) %>%
  mutate(modularity = graph_modularity(group = as.factor(louvain))) %>%
  pull(modularity) %>%
  head(1)

decor_graph %>%  
  get.data.frame(what = "vertices") %>% 
  count(louvain)

nodes <- decor_graph %>% 
  get.data.frame("vertices")
```

@@

The three communities can also be analysed as separate networks. The sites around the Niger Delta are mostly

```{r network 1 centrality in community vs in gen network}
decor_graph %>%
  activate(nodes) %>%
    mutate(deg = centrality_degree(weights = weight)) %>%
  filter(louvain == 1) %>%
  mutate(deg_c = centrality_degree(weights = weight)) %>%
ggraph(layout = "graphopt") + 
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) + 
     scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
 scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = deg, alpha = deg_c), col = "cornflowerblue", show.legend = FALSE) +
  geom_node_text(aes(label = label), col= "black", repel = TRUE) +
      scale_color_manual(values = communitycolor)+
  theme_graph()+
  guides(color = guide_legend(override.aes = list(size = 0.5)), alpha = FALSE, edge_alpha = FALSE, edge_width = guide_legend(override.aes = list(size = 0.3)))+
  ggtitle("network cluster 1 (red)")+
  theme(legend.title = element_text(size = 8, face = "bold"),
              legend.text = element_text(size = 7),
              legend.key.size = unit(0.8, "lines"),
        plot.title = element_text(size = 10, face= "bold"))

decor_graph %>%
  activate(nodes) %>%
    mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 1) %>%
  mutate(deg_c = centrality_eigen(weights = weight)) %>%
flatten(weighted = TRUE) %>% 
ggraph(layout = "graphopt") + 
  geom_edge_fan(aes(edge_width = weight, edge_alpha = weight)) + 
     scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
 scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = deg, alpha = deg_c), col = "cornflowerblue", show.legend = FALSE) +
  geom_node_text(aes(label = label), col= "black", repel = TRUE) +
      scale_color_manual(values = communitycolor)+
  theme_graph()+
  guides(color = guide_legend(override.aes = list(size = 0.5)), alpha = FALSE, edge_alpha = FALSE, edge_width = guide_legend(override.aes = list(size = 0.3)))+
  ggtitle("network cluster 1 (red)")+
  theme(legend.title = element_text(size = 8, face = "bold"),
              legend.text = element_text(size = 7),
              legend.key.size = unit(0.8, "lines"),
        plot.title = element_text(size = 10, face= "bold"))

decor_graph %>%
  activate(nodes) %>%
    mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 1) %>%
ggraph(layout = "graphopt") + 
geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) + 
     scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
 scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(size = 3, colour = "cornflowerblue") +
    geom_node_text(aes(label = label), col= "black", repel = TRUE) +
  theme_graph()

decor_graph %>%
  activate(nodes) %>%
    mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 1) %>%
ggraph(x = lon, y = lat) + 
geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) + 
     scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
 scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(size = 3, colour = "cornflowerblue") +
    geom_node_text(aes(label = label), col= "black", repel = TRUE) +
  theme_graph()
```

```{r network 2 centrality in community vs in gen network}
decor_graph %>%
  activate(nodes) %>%
    mutate(deg = centrality_degree(weights = weight)) %>%
  filter(louvain == 2) %>%
  mutate(deg_c = centrality_degree(weights = weight)) %>%
 activate(edges) %>%
ggraph(layout = "graphopt") + 
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) + 
     scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
 scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = deg, alpha = deg_c), col = "tomato", show.legend = FALSE) +
  geom_node_text(aes(label = label), col= "black", repel = TRUE) +
      scale_color_manual(values = communitycolor)+
  theme_graph()+
  guides(color = guide_legend(override.aes = list(size = 0.5)), alpha = FALSE, edge_alpha = FALSE, edge_width = guide_legend(override.aes = list(size = 0.3)))+
  ggtitle("network cluster 1 (red)")+
  theme(legend.title = element_text(size = 8, face = "bold"),
              legend.text = element_text(size = 7),
              legend.key.size = unit(0.8, "lines"),
        plot.title = element_text(size = 10, face= "bold"))

decor_graph %>%
  activate(nodes) %>%
    mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 2) %>%
  mutate(deg_c = centrality_eigen(weights = weight)) %>%
flatten(weighted = TRUE) %>% 
ggraph(layout = "graphopt") + 
  geom_edge_fan(aes(edge_width = weight, edge_alpha = weight)) + 
     scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
 scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = deg, alpha = deg_c), col = "tomato", show.legend = FALSE) +
  geom_node_text(aes(label = label), col= "black", repel = TRUE) +
      scale_color_manual(values = communitycolor)+
  theme_graph()+
  guides(color = guide_legend(override.aes = list(size = 0.5)), alpha = FALSE, edge_alpha = FALSE, edge_width = guide_legend(override.aes = list(size = 0.3)))+
  ggtitle("network cluster 1 (red)")+
  theme(legend.title = element_text(size = 8, face = "bold"),
              legend.text = element_text(size = 7),
              legend.key.size = unit(0.8, "lines"),
        plot.title = element_text(size = 10, face= "bold"))

decor_graph %>%
  activate(nodes) %>%
    mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 2) %>%
ggraph(layout = "graphopt") + 
geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) + 
     scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
 scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(size = 3, colour = "tomato") +
    geom_node_text(aes(label = label), col= "black", repel = TRUE) +
  theme_graph()

decor_graph %>%
  activate(nodes) %>%
    mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 2) %>%
ggraph(x = lon, y = lat) + 
geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) + 
     scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
 scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(size = 3, colour = "tomato") +
    geom_node_text(aes(label = label), col= "black", repel = TRUE) +
  theme_graph()
```

```{r network 3 centrality in community vs in gen network}
decor_graph %>%
  activate(nodes) %>%
    mutate(deg = centrality_degree(weights = weight)) %>%
  filter(louvain == 3) %>%
  mutate(deg_c = centrality_degree(weights = weight)) %>%
 activate(edges) %>%
ggraph(layout = "graphopt") + 
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) + 
     scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
 scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = deg, alpha = deg_c), col = "darkgreen", show.legend = FALSE) +
  geom_node_text(aes(label = label), col= "black", repel = TRUE) +
      scale_color_manual(values = communitycolor)+
  theme_graph()+
  guides(color = guide_legend(override.aes = list(size = 0.5)), alpha = FALSE, edge_alpha = FALSE, edge_width = guide_legend(override.aes = list(size = 0.3)))+
  ggtitle("network cluster 1 (red)")+
  theme(legend.title = element_text(size = 8, face = "bold"),
              legend.text = element_text(size = 7),
              legend.key.size = unit(0.8, "lines"),
        plot.title = element_text(size = 10, face= "bold"))

decor_graph %>%
  activate(nodes) %>%
    mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 3) %>%
  mutate(deg_c = centrality_eigen(weights = weight)) %>%
flatten(weighted = TRUE) %>% 
ggraph(layout = "graphopt") + 
  geom_edge_fan(aes(edge_width = weight, edge_alpha = weight)) + 
     scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
 scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = deg, alpha = deg_c), col = "darkgreen", show.legend = FALSE) +
  geom_node_text(aes(label = label), col= "black", repel = TRUE) +
      scale_color_manual(values = communitycolor)+
  theme_graph()+
  guides(color = guide_legend(override.aes = list(size = 0.5)), alpha = FALSE, edge_alpha = FALSE, edge_width = guide_legend(override.aes = list(size = 0.3)))+
  ggtitle("network cluster 1 (red)")+
  theme(legend.title = element_text(size = 8, face = "bold"),
              legend.text = element_text(size = 7),
              legend.key.size = unit(0.8, "lines"),
        plot.title = element_text(size = 10, face= "bold"))

decor_graph %>%
  activate(nodes) %>%
    mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 3) %>%
ggraph(layout = "graphopt") + 
geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) + 
     scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
 scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(size = 3, colour = "darkgreen") +
    geom_node_text(aes(label = label), col= "black", repel = TRUE) +
  theme_graph()

decor_graph %>%
  activate(nodes) %>%
    mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 3) %>%
ggraph(x = lon, y = lat) + 
geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) + 
     scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
 scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(size = 3, colour = "darkgreen") +
    geom_node_text(aes(label = label), col= "black", repel = TRUE) +
  theme_graph()
```

## dont use this
```{r relative edge weights and centralization by cluster}
# Relative edge weights (how important are the connections within the cluster)
rel_edge_weight <- vector("double", length(3))
for (i in 1:3) {
  rel_edge_weight[[i]] <- decor_graph %>%
    # filter community
    activate(nodes) %>%
    filter(louvain == i) %>%
    # relative edge weight (sum of weights / number of edges)
    activate(edges) %>%
    as.data.frame() %>%
    {
      sum(.$weight) / nrow(.)
    }
}

@@

## checking the centrality of the nodes of a cluster

centrality_deg.cluster <- decor_graph %>%
    # filter community
    activate(nodes) %>%
    #filter(louvain == i) %>%
    # degree centralization
  #  mutate(cent = centrality_degree("all", normalized = TRUE)) %>%
  mutate(centr = centrality_degree(weights = weight, mode = "all", normalized = TRUE)) %>% 
#centr_degree(mode = "all", normalized = TRUE, weights = weight) %>%
    as_tibble()

centrality_eigen.cluster <- decor_graph %>%
    # filter community
    activate(nodes) %>%
    #filter(louvain == i) %>%
    # degree centralization
  #  mutate(cent = centrality_degree("all", normalized = TRUE)) %>%
  mutate(centr = centrality_eigen(weights = weight)) %>% 
#centr_degree(mode = "all", normalized = TRUE, weights = weight) %>%
    as_tibble()

centr_deg.comb.cluster <- c()
for (i in 1:3) {

  centr_deg.comb.cluster[i] <- centrality_deg.cluster %>%
    # filter community
    filter(louvain == i) %>%
    # degree centralization
    summarise(centr_comb = mean(centr)) %>%
    as_tibble() %>% 
    pull(centr_comb)

}

centr_eig.comb.cluster <- c()
for (i in 1:3) {

  centr_eig.comb.cluster[i] <- centrality_eigen.cluster %>%
    # filter community
    filter(louvain == i) %>%
    # degree centralization
    summarise(centr_comb = mean(centr)) %>%
    as_tibble() %>% 
    pull(centr_comb)

}

# put together
tibble(rel_edge_weight, 
       centralization,
    centr_deg.comb.cluster, 
     centr_eig.comb.cluster
     ) %>%
  mutate(cluster = 1:3) %>%
  mutate(rel_edge_weight = rel_edge_weight / 100,
        centr_deg.comb.cluster = centr_deg.comb.cluster/1000,
         centr_eig.comb.cluster = centr_eig.comb.cluster/0.1
        ) %>%
  rename("relative edge weight" = rel_edge_weight) %>%
  pivot_longer(!cluster, names_to = "key", values_to = "value") %>%
  # plot
  ggplot(aes(x = cluster, y = value, fill = key)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_grey() +
  theme(axis.text.x = element_text(angle = 90)) +
  theme_publish()
```

## dont use this

```{r relative edge weights, roles and centralization by cluster as boxplot}
# Relative edge weights (how important are the connections within the cluster)
rel_edge_weight <- data.frame()
#for (i in 1:3) {
  rel_edge_weight <- decor_graph %>%
    # filter community
    activate(nodes) %>%
  #  filter(louvain == i) %>%
    # relative edge weight (sum of weights / number of edges)
    activate(edges) %>%
    as.data.frame() %>%
   mutate(louvain = i,
          parameter = "edge_weight") %>% 
    rename(value = weight) %>% 
  select(value, louvain, parameter)
#}

centralization <- data.frame(value = centralization, louvain = c(1:3), parameter = "centralization")
## checking the centrality of the nodes of a cluster

centrality_deg.cluster <- decor_graph %>%
    # filter community
    activate(nodes) %>%
    #filter(louvain == i) %>%
    # degree centralization
  #  mutate(cent = centrality_degree("all", normalized = TRUE)) %>%
  mutate(centr = centrality_degree(weights = weight, mode = "all", normalized = TRUE)) %>% 
#centr_degree(mode = "all", normalized = TRUE, weights = weight) %>%
    as_tibble()

centrality_eigen.cluster <- decor_graph %>%
    # filter community
    activate(nodes) %>%
    #filter(louvain == i) %>%
    # degree centralization
  #  mutate(cent = centrality_degree("all", normalized = TRUE)) %>%
  mutate(centr = centrality_eigen(weights = weight)) %>% 
#centr_degree(mode = "all", normalized = TRUE, weights = weight) %>%
    as_tibble()

centr_deg.comb.cluster <- c()
for (i in 1:3) {

  centr_deg.comb.cluster[i] <- centrality_deg.cluster %>%
    # filter community
    filter(louvain == i) %>%
    # degree centralization
    summarise(centr_comb = mean(centr)) %>%
    as_tibble() %>% 
    pull(centr_comb)

}

centr_eig.comb.cluster <- c()
for (i in 1:3) {

  centr_eig.comb.cluster[i] <- centrality_eigen.cluster %>%
    # filter community
    filter(louvain == i) %>%
    # degree centralization
    summarise(centr_comb = mean(centr)) %>%
    as_tibble() %>% 
    pull(centr_comb)

}

# put together
tibble(rel_edge_weight, 
     #  centralization,
      #centr_deg.comb.cluster, 
     #  centr_eig.comb.cluster
     ) %>%
  mutate(cluster = 1:3) %>%
  mutate(rel_edge_weight = rel_edge_weight / 100,
        centr_deg.comb.cluster = centr_deg.comb.cluster/1000#,
        #  centr_eig.comb.cluster = centr_eig.comb.cluster/0.1
        ) %>%
  rename("relative edge weight" = rel_edge_weight) %>%
  pivot_longer(!cluster, names_to = "key", values_to = "value") %>%
  # plot
  ggplot(aes(x = cluster, y = value, fill = key)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_grey() +
  theme(axis.text.x = element_text(angle = 90)) +
  theme_publish()
```

```{r add community to edges}

#mark which edges are in which community or without
decor_graph <- decor_graph %>% 
  activate(edges) %>% 
  mutate(community = case_when(from %in% which(nodes$louvain == 1) & to %in% which(nodes$louvain == 1) ~ "1",
                               from %in% which(nodes$louvain == 2) & to %in% which(nodes$louvain == 2) ~ "2",
                               from %in% which(nodes$louvain == 3) & to %in% which(nodes$louvain == 3) ~ "3",
         TRUE ~ "crossing"))

```{r}
decor_graph %>% 
  get.data.frame(what = "edges") %>% 
  count(community)

edge_inside <- decor_graph %>% 
  activate(edges) %>% 
  filter(community != "crossing") %>% 
  activate(nodes) %>% 
  mutate(centr_comm = centrality_degree(weights = weight, mode = "all")) %>% 
  get.data.frame("vertices") %>% 
  mutate(centr_comm_weigthed = case_when(louvain == 1 ~ centr_comm/8,
                                         louvain == 2 ~ centr_comm/6,
                                         TRUE ~ centr_comm/3))%>% 
  mutate(type = "inside")

edge_outside <- decor_graph %>% 
  activate(edges) %>% 
  filter(community == "crossing") %>% 
  activate(nodes) %>% 
  mutate(centr_comm = centrality_degree(weights = weight, mode = "all")) %>% 
  get.data.frame("vertices") %>% 
  mutate(centr_comm_weigthed = case_when(louvain == 1 ~ centr_comm/(19-9),
                                         louvain == 2 ~ centr_comm/(19-7),
                                         TRUE ~ centr_comm/(19-4))) %>% 
  mutate(type = "outside")

data <- rbind(edge_inside, edge_outside) %>% 
  mutate(louvain = as.character(louvain))

# plot as a boxplot
ggplot(data, aes(x = louvain, y = centr_comm_weigthed, fill = type)) +
  geom_boxplot() +
  scale_fill_manual(values = c("grey80", "black")) +
  xlab("Community") +
  ylab("Total Edge Weight") +
  ggtitle("Weights of Edges Within and Between Communities")

ggplot(data, aes(x = louvain, y = centr_comm, fill = type)) +
  geom_boxplot() +
  scale_fill_manual(values = c("grey80", "black")) +
  xlab("Community") +
  ylab("Total Edge Weight") +
  ggtitle("Weights of Edges Within and Between Communities")
```

```{r}
data <- rbind(edge_inside, edge_outside) %>% 
  mutate(louvain_bb = as.character(louvain))

data %>% 
  group_by(louvain, type) %>% 
  summarise(centr_comm = mean(centr_comm),
            centr_comm_weigthed = mean(centr_comm_weigthed)) %>%
  ggplot(aes(x = louvain, y = centr_comm_weigthed, fill = type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Community", y = "Mean centrality") +
  scale_fill_manual(values = c("red", "green", "blue")) +
  theme_classic()

data %>% 
  group_by(louvain, type) %>% 
  summarise(centr_comm_weigthed = mean(centr_comm_weigthed)) %>%
  pivot_wider(names_from = type, values_from = centr_comm_weigthed) %>%
  mutate(ratio_in_out = inside / outside)

```{r edge doenst work}
library(dplyr)
library(ggplot2)

# group edges by community and whether they are within or outside the community
edge_groups <- decor_graph %>% 
  activate(edges) %>%
  mutate(
    community = ifelse(
      from %in% nodes$louvain & to %in% nodes$louvain & nodes$louvain[from] == nodes$louvain[to],
      as.character(nodes$louvain[from]),
      "Crossing"
    ),
    is_within = ifelse(community == "Crossing", "Crossing", "Within")
  ) %>%
  group_by(community, is_within) %>%
  summarize(weight = sum(weight))

# create boxplot
ggplot(edge_groups, aes(x = community, y = weight, fill = is_within)) +
  geom_boxplot() +
  labs(title = "Weight of Edges Within and Outside Communities",
       x = "Community",
       y = "Edge Weight")

```{r trash }
# plot the weight of edges within the community
ggplot(edges_within_community, aes(x = in_community, y = weight)) +
  geom_boxplot() +
  xlab("Community") +
  ylab("Edge weight")

# calculate edge weights
g <- g %>% activate(edges) %>% mutate(weight = sample(1:10, n(), replace = TRUE))

g <- decor_graph

# calculate total weights within each community
in_community <- g %>% 
  activate(edges) %>%
  group_by(from, to, louvain) %>%
  summarise(total_weight = sum(weight)) %>%
  ungroup()

# calculate total weights between communities
between_communities <- g %>% 
  activate(edges) %>%
  filter(!from %--% to) %>%
  group_by(from, to) %>%
  summarise(total_weight = sum(weight)) %>%
  ungroup()

# combine the data frames for plotting
data <- rbind(in_community, between_communities)

# plot as a boxplot
ggplot(data, aes(x = community, y = total_weight, fill = from %--% to)) +
  geom_boxplot() +
  scale_fill_manual(values = c("grey80", "black")) +
  xlab("Community") +
  ylab("Total Edge Weight") +
  ggtitle("Weights of Edges Within and Between Communities")

library(tidygraph)
library(igraph)

# Create a tidygraph object from your data

# Calculate the communities using the Louvain algorithm
decor_graph <- decor_graph %>% activate(nodes) %>% mutate(louvain = group_louvain())

# Mark edges that are within a community
#decor_graph <- decor_graph %>% 
#  activate(edges) %>% 
  #mutate(in_community = group_edges(decor_graph, "louvain", .f = function(x) all(x == x[1])))

# `in_community` column will be `TRUE` for edges within a community and `FALSE` for edges crossing communities
decor_graph <- decor_graph %>%
  activate(edges) %>%
  mutate(
    community = ifelse(
      from %in% nodes$louvain & to %in% nodes$louvain & nodes$louvain[from] == nodes$louvain[to],
      as.character(nodes$louvain[from]),
      "crossing"
    )
  )

@@

  # find the edges that cross between communities
  mutate(crossing = community_crossing(weights = weight)) %>%
  filter(
    crossing == TRUE,
    weight >= 10
  ) %>%
  activate(nodes) %>% 
  filter(!node_is_isolated()) %>%

@@

new cluster comparison by centralisation and centrality

```{r cluster centralisation, fig.cap="Degree and eigen centralization by cluster"}
### Centralization

# Function to calculate centralization for each cluster
calculate_centralization <- function(graph, louvain_column) {
  # Create an empty tibble to store the centralization results
  centralization <- tibble(cluster = character(), cent_deg = double(), cent_eigen = double())

  # Get the unique clusters from the specified louvain column
  unique_clusters <- graph |>
    pull({{louvain_column}}) |>
    unique()

  # Loop through each cluster
  for (cluster in unique_clusters) {
    # Filter the graph based on the current cluster
    subgraph <- graph |>
      filter({{louvain_column}} == cluster)

    # Calculate the centralization measures for the subgraph
    cent_deg <- centr_degree(subgraph, normalized = TRUE)$centralization
    cent_eigen <- centr_eigen(subgraph, normalized = TRUE)$centralization

    # Add the results to the centralization tibble
    centralization <- centralization |>
      add_row(cluster = as.character(cluster), cent_deg = cent_deg, cent_eigen = cent_eigen)
  }

  # Arrange the centralization results by cluster and return the tibble
  centralization <- centralization |>
    arrange(cluster)
  return(centralization)
}

# Calculate centralization for the given graph and louvain column
calculate_centralization(decor_graph, louvain)

calculate_centralization(decor_graph, louvain) |> 
  pivot_longer(-cluster, names_to = "cent_type") |> 
  ggplot() +
  geom_bar(aes(x = cent_type, y = value, fill = cluster), stat = "identity", position = "dodge") +
  theme_minimal() +
  labs(x = "Centrality Type", y = "Centrality Value", fill = "Cluster")
```

```{r cluster centralities, fig.cap="Mean centralities (Degree & Eigen) by cluster"}
### Centralities

# Function to calculate centralities for each cluster group
pull_centralities_clust <- function(graph) {
  # Activate the nodes in the graph
  graph |>
    activate(nodes) |>
    # Calculate degree centrality and eigenvector centrality
    mutate(deg = centrality_degree(weights = weight, normalized = TRUE), eig = centrality_eigen(weights = weight)) |>
    # Select the louvain cluster, degree centrality, and eigenvector centrality columns
    select(label,louvain, deg, eig) |>
    # Convert the result to a tibble
    as_tibble()
}

# Calculate centralities for each cluster
centr <- pull_centralities_clust(decor_graph)

# Display boxplots of centralities for degree centrality
a <- ggplot(centr) +
  geom_boxplot(aes(x = louvain, y = deg, fill = louvain)) +
  scale_fill_grey(start = 0.5, end  = 1) +
  theme_bw() +
  theme(legend.position = "none")+
  labs(y = "degree centrality",
       x = "community")

# Display boxplots of centralities for eigenvector centrality
b <- ggplot(centr) +
  geom_boxplot(aes(x = louvain, y = eig, fill = louvain)) +
  scale_fill_grey(start = 0.5, end  = 1) +
  theme_bw() +
  theme(legend.position = "none")+
  labs(y = "eigenvector centrality",
       x = "community")

# Combine the boxplots
a + b

# Summary of centralities per cluster
centr |>
  group_by(louvain) |>
  summarise(meandeg = mean(deg), meaneig = mean(eig))
```

If we take a closer look at the "internal" connections of community 1, we see that "braided_strip_roulette" is present at all sites. The other main decor "straight_cord_mat" on the other hand is missing on 2 sites: Oudalan and Prom. The two decors are spread across the whole area of the community. The only other decor that is common in the community is the "twisted_cord_roulette" which is also abundant across the whole research area. It makes up a rather large part of the assemblage in Oudalan and Prom but was not found at the sites of Sadia and Ambere Dougon. The fact that SAO has a high degree of centrality shows that the location is not decisive for the connectedness within this community. This is underlined by the fact that the less central sites are distributed across the whole area. There is an ambivalence between a relative consistency across the whole area and the diversity between sites that are close to each other. One reason for this phenom can be the contact to other communities in the mountain area. Damassogou exemplifies this as the site contains the defining decors of community 1 and 2. But the difference between Oudalan and SAO that are geographically isolated suggest that there might be more reasons for this diversity. Community 2 is less diverse with "braided_cord_roulette" as the only outstanding decor type. The weaker connections of the sites Kokolo and Damassogou are based on data bias (the proportion is counted in regard to all sherds not only the decorated ones), the decor was probably equally abundant there. Other decors spread across the community are "folded_strip_roulette" and "dragged_comb". The first one is missing in KAN 4 and TOD H while the second one is missing at Kokolo and Damassogou. The other indicator for the homogeneous character of the community is the nearly equal degree of centrality of every site. As mentioned above community 3 is more diverse but also has weaker connections. Instead of one defining decor there are several ones but none is abundant across the whole community. The spatial distribution differs between the decors. While "paint" and "twisted_cord_roulette" appear from Gao to TOG 2, "folded_strip_roulette", "channel" and "incised" are located in the western part of the community's area. Essouk and TMD have the highest degree of centrality. Even though it is located at the margin of the research area, its character as a "market"-site explains the amount of links it has. The role of TMD for the network is elaborated elsewhere (!).

@@

## Simplified network backbone

@@

```{r backbone extraction}
# extract binary backbones at decreasing levels of significance
decor.05 <- extract_backbone(decor.flat, 0.05)
decor.15 <- extract_backbone(decor.flat, 0.15)
decor.20 <- extract_backbone(decor.flat, 0.18)

@@

```{r import early sites}
library(readr)
decorsearly <- read_csv2(here::here("analysis/data/raw_data/decors_early.csv")) %>% 
  select(-1)

#decorsearly <- decorsearly%>%
#  filter(site != "SAO 3" & site != "Oudalan Fer moyen")  %>%
 ## mutate_at(vars(-site), as.numeric) %>% 
 # mutate(site = case_when(site == "Akumbu 1" ~ "Akumbu early",
     #    site == "TMD" ~ "TMD 3",
   #      TRUE ~ site))

#write.csv2(decorsearly, here::here("analysis/data/raw_data/decors_early.csv"))
# apply function to transform frequency table into edge list
edgelist <- make_edgelist(decorsearly)

@@

# Sparsified clustering

## late cluster
### calculating cluster with sparsification

here we test clustering a sparsified network in contrast to the "normal" network. first we have to get the new clusters and then add them as louvain_bb to our decor_graph
```{r}
decor.20 <- decor.20 %>%
  activate(nodes) %>%
  mutate(louvain = as.character(group_louvain()))

# calculate modularity of louvain clustering
modularity.20 <- decor.20 %>%
  activate(nodes) %>%
  mutate(modularity = graph_modularity(group = as.factor(louvain))) %>%
  pull(modularity) %>%
  head(1)

decor.20 %>%  
  get.data.frame(what = "vertices") %>% 
  count(louvain)

decor_graph <- decor_graph %>% 
    activate(nodes) %>%
  mutate(louvain_bb = as.character(decor.20 %>% 
  get.data.frame("vertices") %>% 
  pull(louvain)))

nodes <- decor_graph %>% 
  get.data.frame("vertices")

### mapping the new clusters

ggraph(decor_graph, x = lon, y = lat) +
  geom_sf(data = mali_adm0, fill = NA) +
  geom_sf(data = rivers, colour = "black") +
  edges_community +
  ggforce::geom_mark_hull(aes(x = lon, y = lat, fill = louvain_bb, colour = louvain_bb),
                          expand = unit(1.5, "mm"),
                          radius = unit(1.5, "mm")) +
  scale_fill_manual(values = communitycolor)+
  scale_colour_manual(values = communitycolor)+
  nodes_community +
  theme_community +
  coord_sf(xlim = c(-7, 2), ylim = c(13, 19), expand = FALSE) +
  theme_graph()
```

### taking a look at all 3 clusters

he three communities can also be analysed as separate networks. The sites around the Niger Delta are mostly

```{r network 1 centrality in community vs in gen network}
